---
title: "Introduction to the RealHome Price package"
author: "Nicholas scholl,Russel Marvin, Bipin Parajuli, Moreen Owino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the RealHome Price package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
install.packages("RealHomePrice")
devtools::document()
library(RealHomePrice)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(neuralnet)
library(sigmoid)
library(broom)
library(caTools)
library(devtools)
library(usethis)
```


### Introduction

This vignette provides an introduction to the `RealHomePrice` package for estimating house prices.The package offers a general estimation of the price that a house could go for based on the values of three variables which are:

- the number of bedrooms in the house,
- the number of bathrooms 
- the house size.

This package is important when trying to figure out whether a property is overpriced or within acceptable market prices.

There are currently no packages that attempt to predict the price of a house and this is the first of its kind.The package uses a neural network model and linear model trained on an existing real estate data set that come with the package and are used to predict the prices.

Existing resources in R have only used either linear regression or the Arima model to predict the actual housing prices and our package makes an attempt in using a neural network to do this.

### Data sets

*realtor_data.csv* This data set was used in training both the linear model and the neural network.The data was cleaned and the required variables were selected and used to train the models.

*tests.rds* <- This rds file contains the data on the predicted and actual price values obtained from doing tests on the neural net model.

### Models

The package comes with pre loaded models obtained from training using the train data.

*pricemodel.rds* contains the neural network model
*linear_model_fit.rds* contains the linear model

### functions

The package starts by taking input from the user by asking them to key in values for the three house variables i.e number of bedrooms:bed, number of bathrooms:bath and the house size in square feet and creates a data frame of the variables as shown in the code below

*Taking user input*

```{r}

{
    bed = as.numeric(readline("Enter bed : "));
    bath = as.numeric(readline("Enter bath : "));
    house_size = as.numeric(readline("Enter house_size: "));
}
user_data = data.frame(bed, bath, house_size)
user_data
```

*normalize_input function*

```{r}
normalize_input <- function(user_Data) {
  bed = (user_Data[1,1]-min(cleaned_realtordata$bed))/(max(cleaned_realtordata$bed))
  bath = (user_Data[1,2]-min(cleaned_realtordata$bath))/(max(cleaned_realtordata$bath))
  house_size = (user_Data[1,3]-min(cleaned_realtordata$house_size))/(max(cleaned_realtordata$house_size))
  nor_User_data <- data.frame(bed, bath, house_size)
  
  return(nor_User_data)
}

```


The normalize_input function takes the user data data frame as an argument and returns the normalized values for each of the variables by subtracting the minimum value of the variable and dividing by the maximum value of the variable as given in the cleaned realtor_data.csv.The output of the function on user data is as shown below.

```{r}
nor_User_data <- normalize_input(user_data)
nor_User_data
```

*predict_lm function*

The predict_lm function reads in the  the linear model RDS that comes with the package and uses the model to predict the scaled house price based on the normalized user data

```{r}
predict_lm <- function(nor_User_data) {
  if (class(linear_model_fit) != "lm"
  ) stop("This function needs an LM model")
  predict(linear_model_fit, nor_User_data)
}

pricelm <-predict_lm(nor_User_data)
pricelm
```


*unscale_price function*

The unscale_price function is the used to de normalize the scaled price predicted by the models back to their original real values by doing carrying out the reverse of what had been done to normalize it. It therefore multiplies the price value by the maximum and adds back the minimum value for the price variable.

```{r}
unscale_price <- function(price) {
  price*(max(cleaned_realtordata$price))+min(cleaned_realtordata$price)
}

# for example

lm_price<-unscale_price(pricelm)
lm_price
```

*predict_nn function*

```{r}
PriceModel <- function(nor_User_data) {
   
  predict(neural_net_fit, nor_User_data)
}
```

This function takes in the normalized user data as an argument and reads in the trained neuralnetwork model that comes with the package and uses the model to predict the scaled/normalized value of the house price.This scaled output will then be unscaled using the unscale_price function mentioned above to get the real actual house price.

*predict_price function*

Finally we have the predict function that calls all of the functions above and 
returns the unscaled predicted house prices for both the linear model and neural network model.

```{r}
predict_Price <- function(user_data) {
  nor_User_data <- normalize_input(user_data)
  lm_price <- predict_lm(nor_User_data)
  nn_price <- PriceModel((nor_User_data))
  lm_price <- unscale_price(lm_price)
  nn_price <- unscale_price(nn_price)
  prices <- data.frame(lm_price, nn_price)
  return (prices)
 
}
```

The predict_price function in use as shown below:

```{r}

user_data = data.frame(bed, bath, house_size)
predict_Price(user_data)
```
*Plotting the neural net model*

The neuralnetwork model can be plotted by reading in the RDS containing the model and using the plot function to plot the model

```{r}

#plotpoint <- load("PriceModel.rdata")
#plot(plotpoint)
```

*plotting the models predicted prices vs actual prices obtained from running  test data through the neural network model*

The package comes loaded with the tests.rds data set that contains actual price and predicted price data that can be loaded using the readRDS function and  plotted using ggplot to get an idea of the model efficiency as shown below:
```{r}
test_results<- readRDS("tests.rds")
test_results
```

```{r}
ggplot(test_results,aes(x = price, y = predictedprice))+geom_point()
```


The plot above shows that there is indeed some correlation between the actual price and the price predicted by the model and thus the model can be used to estimate house prices.

There might be a slight difference between actual and predicted prices due to other factors such as location of the house ,acre lot etc which where not included in the training but the model can provide an estimate of how much the house could cost.




