---
title: "Introduction to the RealHomePrice package"
author: "Nicholas scholl,Russel Marvin, Bipin Parajuli, Moreen Owino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the RealHome Price package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
#library(RealHomePrice)
library(ggplot2)
```


### Introduction

This vignette provides an introduction to the `RealHomePrice` package used for estimating house prices.The package offers a general estimation of the price that a house could go for based on the values of three variables which are:

- the number of bedrooms in the house,
- the number of bathrooms 
- the house size.

This package is important when trying to figure out whether a property is overpriced or within acceptable market prices.

There are currently no packages that attempt to predict the actual price of a house and this is the first of its kind.The package uses a neural network model and linear model trained on an existing real estate data set and this models come with the package and are used to predict the prices.

Existing resources in R have only used either linear regression or the Arima model to predict the actual housing prices and our package makes an attempt in using a neural network to do this.

### Data sets

*cleaned_realtor.rdata* - This data set was used in training both the linear model and the neural network.The data was cleaned by removing all missing values and outliers and the required variables i.e bed,bath and house size were selected and used to train the models.

*tests2.RData* - This file contains the scaled data on the predicted and actual price values obtained from doing tests on the neural net model.

*train_data.rdata* - This file contains the a subset of the cleaned_realtor data that was normalized and used to train the models.

*lm_model_test.RData* - It contains the actual and predicted prices on running test data on the linear model.

*nn_model_test.RData* - It contains the actual and predicted prices on running test data on the neuralnet model.

### Models

The package comes with a pre loaded neural net model obtained from training using the train data.

*neuralmodel3.RData* contains the neural network model obtained from running the following function : neuralmodel <- neuralnet(price~bed+bath+house_size,data=train_data,hidden=2, threshold=0.01,act.fct=relu, rep=1, linear.output=TRUE)


### functions


*loading the package*

The package can be installed by running the following codes

```{r}

#install.packages("RealHomePrice")
# then load the library
#library(RealHomePrice)
```


The package starts by taking input from the user by asking them to key in values for the three house variables i.e number of bedrooms:bed, number of bathrooms:bath and the house size in square feet and creates a data frame of the variables as shown in the code below

*Taking user input*

```{r}

{
    #bed = as.numeric(readline("Enter bed : "));
    #bath = as.numeric(readline("Enter bath : "));
    #house_size = as.numeric(readline("Enter house_size: "));
}
#user_data = data.frame(bed, bath, house_size)
#user_data
```


```{r}

bed = 3
bath = 2
house_size = 1000

user_data = data.frame(bed, bath, house_size)

user_data


```



*normalize_input function*

```{r}

load("C:/RealHomePrice/data/cleaned_realtordata.rdata")

normalize_input <- function(user_Data) {

  bed = (user_Data[1,1]-min(cleaned_realtordata$bed))/(max(cleaned_realtordata$bed))
  bath = (user_Data[1,2]-min(cleaned_realtordata$bath))/(max(cleaned_realtordata$bath))
  house_size = (user_Data[1,3]-min(cleaned_realtordata$house_size))/(max(cleaned_realtordata$house_size))
  nor_User_data <- data.frame(bed, bath, house_size)
  
  return(nor_User_data)
}

```


The normalize_input function takes the user data data frame as an argument and returns the normalized values for each of the variables by subtracting the minimum value of the variable and dividing by the maximum value of the variable as given in the cleaned realtor_data.The output of the function on user data is as shown below.

```{r}
nor_User_data <- normalize_input(user_data)
nor_User_data
```

*predict_lm function*

The predict_lm function  uses the linear model that comes loaded in the package to predict the scaled house price based on the normalized user data(nor_user_data) obtained from the normalize_input function above.


```{r}

load("C:/RealHomePrice/data/Linear_model_fit.Rdata")


```




```{r}

predict_lm <- function(nor_User_data, linear_model_fit1 = linear_model_fit) {
  if (class(linear_model_fit) != "lm"
  ) stop("This function needs an LM model")
  predict(linear_model_fit, nor_User_data)
}

# for example using the nor_user_data we got from the normalized input function.
pricelm <-predict_lm(nor_User_data)
pricelm
```


*unscale_price function*

The unscale_price function is the used to de normalize the scaled price predicted by the models back to their original real values by doing carrying out the reverse of what had been done to normalize it. It therefore multiplies the price value by the maximum and adds back the minimum value for the price variable.

```{r}
unscale_price <- function(price) {
  price*(max(cleaned_realtordata$price))+min(cleaned_realtordata$price)
}

# for example

lm_price<-unscale_price(pricelm)
lm_price
```

*predict_nn function*

```{r}
predict_nn <- function(nor_User_data) {
  load("C:/RealHomePrice/data/neuralmodel3.RDa")
  predict(neuralmodel, nor_User_data)
}

# example
price<- predict_nn(nor_User_data)
price
```

This function takes in the normalized user data as an argument and reads in the trained neuralnetwork model that comes with the package and uses the model to predict the scaled/normalized value of the house price.This scaled output will then be unscaled using the unscale_price function mentioned above to get the real actual house price.

*predict_price function*

Finally we have the predict function that calls all of the functions above and 
returns the unscaled predicted house prices for both the linear model and neural network model.

```{r}
predict_Price <- function(user_data) {
  nor_User_data <- normalize_input(user_data)
  lm_price <- predict_lm(nor_User_data)
  nn_price <- predict_nn((nor_User_data))
  lm_price <- unscale_price(lm_price)
  nn_price <- unscale_price(nn_price)
  prices <- data.frame(lm_price, nn_price)
  return (prices)
 
}
```

The predict_price function in use as shown below:

```{r}
{
    bed = as.numeric(readline("Enter bed : "));
    bath = as.numeric(readline("Enter bath : "));
    house_size = as.numeric(readline("Enter house_size: "));
}
user_data = data.frame(bed, bath, house_size)
predict_Price(user_data)
```
*Plotting the neural net model*

The neuralnetwork model can be plotted by loading in the neuralmodel3.RData file which contains the trained neuralmodel and using the plot function to plot the model

```{r}
load("C:/RealHomePrice/data/neuralmodel3.Rda")

plot(neuralmodel)

```
The plot above shows that the model used two hidden layers and had a low error  value of 0.059695 showing that the model had a relatively good performance on the data set since acceptable error values are usually between 0 and 1.


*plotting the models scaled predicted prices vs actual prices obtained from running  test data through the neural network model*

The package comes loaded with the tests2.RData data set that contains actual prices and predicted prices data that can be loaded using the load function and  plotted using ggplot to get an idea of the model efficiency as shown below:
```{r}
load("C:/RealHomePrice/data/tests2.RData")
```

```{r}

actual_predicted_prices |> ggplot(aes(x =Actualprice, y =nn_predictedprice))+ geom_point() + 
  geom_smooth(method = 'lm')+
  ggtitle("scaled predicted vs real prices")

```


The plot above shows that there is indeed some correlation between the actual price and the price predicted by the model and thus the model can be used to estimate house prices.Most of the points are close to the line of best fit.

There might be a slight difference between actual and predicted prices due to other factors such as location of the house ,acre lot etc which where not included in the training but the model can provide an estimate of how much the house could cost.

*Plotting real Vs Actual prices obtained from running the test data through the linear model and the neuralnet model*
First we load the required data sets that were obtained from running the test data

```{r}
load("C:/RealHomePrice/data/nn_model_test.RData")
load("C:/RealHomePrice/data/lm_model_test.RData")
```


```{r}
par(mfrow=c(1,2))
plot(nn_model_test$Actualprice,nn_model_test$nn_predictedprice,col='red',main='Real vs predicted NN',pch=18,cex=0.7)
abline(0,1,lwd=2)
legend('bottomright',legend='NN',pch=18,col='red', bty='n')
plot(lm_model_test$Actualprice,lm_model_test$lm_predictedprices,col='blue',main='Real vs predicted lm',pch=18, cex=0.7)
abline(0,1,lwd=2)
legend('bottomright',legend='LM',pch=18,col='blue', bty='n', cex=.95)
```
Again the plots show that most of the predicted prices can be found close to the best fit line and this shows that the models can be used to predict the house prices.

